### 2009-07-29 SC09/abstract.tex
got the graphs looking good. got the abstract right. got the keywords, crap, authors, copyright.
git-svn-id: https://plfs.svn.sourceforge.net/svnroot/plfs/doc@7 979fb63a-54b4-4514-8d67-9ab8ea1b59df
attached is the most recent version of the paper.  I'll do another big push tomorrow and then I'll submit it since I'll be on vacation all next week.  I think I've addressed most of the reviewer's comments but here's a couple that maybe still need to be addressed (or ignored):
the addition of metadata benchmark such as mdtest would make paper
stronger

the paper does not mention about file locks.  f

\vspace{-.5cm}
\abstract{
\scribble{Rewordings. Eliminated last paragraph. Changed definition of \\upfs to use 'file system' instead of 'file system'. Changed National Science Foundation site to PSC}
\em{
Parallel applications running across thousands of processors must protect
themselves from inevitable component failures. Many applications 
insulate themselves from failures by checkpointing, a
process in which they save their state to persistent storage. Following a
failure, they can resume computation using this state.
For many applications, saving this state into a shared single file
is most convenient.  With such an approach, the size of writes are often 
small and not aligned with file system boundaries. Unfortunately for these
applications, this preferred data layout results in pathologically poor
performance from the underlying file system which is optimized for
large, aligned writes to non-shared files.
\if 0
% Milo's changes which I didn't incorporate
process in which they periodically save their state to persistent storage. Following a
failure, they can retrieve the most recently saved state and resume computation from that
point. For many applications, saving this state into a shared single file, often
formatted for storage by complex libraries such as NetCDF or HDF5,
is most convenient.  With such an approach, the size of individual write system calls are often 
small and not aligned with file system boundaries especially when the application cannot statically define the order of state variables, as with adaptive mesh refinement app. Unfortunately for these
applications, this write access pattern can result in pathologically poor
performance from an underlying file system, itself often optimized for
\fi

@@

We expect that \plfs\ can improve the checkpoint bandwidth for any large
parallel application that writes to a single file. The expected improvement is
especially large for those applications doing unaligned or random IO, patterns
which have become increasingly prevalent recently due to the wide-spread
adoption of complex formatting libraries such as NetCDF and HDF5.

\if 0
% here's the 150 word limited version we submitted to SC
Parallel applications running across thousands of processors must protect themselves from inevitable system failures. Many applications insulate themselves from failures by checkpointing. For many applications, checkpointing into a shared single file is most convenient. With such an approach, the size of writes are often small and not aligned with file system boundaries. Unfortunately for these applications, this preferred data layout results in pathologically poor performance from the underlying file system which is optimized for large, aligned writes to non-shared files. To address this fundamental mismatch, we have developed a virtual parallel log structured file system, PLFS. PLFS remaps an application’s preferred data layout into one which is optimized for the underlying file system. Through testing on PanFS, Lustre, and GPFS, we have seen that this layer of indirection and reorganization can reduce checkpoint time by an order of magnitude for several important benchmarks and real applications.
\fi
}
}

>>>>

\vspace{-.5cm}
\abstract{
\scribble{Rewordings. Eliminated last paragraph. Changed definition of \\upfs to use 'file system' instead of 'file system'. Changed National Science Foundation site to PSC}
\em{
\if 0
Parallel applications running across thousands of processors must protect
themselves from inevitable component failures. Many applications 
insulate themselves from failures by checkpointing, a
process in which they save their state to persistent storage. Following a
failure, they can resume computation using this state.
For many applications, saving this state into a shared single file
is most convenient.  With such an approach, the size of writes are often 
small and not aligned with file system boundaries. Unfortunately for these
applications, this preferred data layout results in pathologically poor
performance from the underlying file system which is optimized for
large, aligned writes to non-shared files.
\if 0
% Milo's changes which I didn't incorporate
process in which they periodically save their state to persistent storage. Following a
failure, they can retrieve the most recently saved state and resume computation from that
point. For many applications, saving this state into a shared single file, often
formatted for storage by complex libraries such as NetCDF or HDF5,
is most convenient.  With such an approach, the size of individual write system calls are often 
small and not aligned with file system boundaries especially when the application cannot statically define the order of state variables, as with adaptive mesh refinement app. Unfortunately for these
applications, this write access pattern can result in pathologically poor
performance from an underlying file system, itself often optimized for
\fi

@@

We expect that \plfs\ can improve the checkpoint bandwidth for any large
parallel application that writes to a single file. The expected improvement is
especially large for those applications doing unaligned or random IO, patterns
which have become increasingly prevalent recently due to the wide-spread
adoption of complex formatting libraries such as NetCDF and HDF5. 
\fi

% here's the 150 word limited version we submitted to SC
Parallel applications running across thousands of processors must protect
themselves from inevitable system failures. Many applications insulate
themselves from failures by checkpointing. For many applications, checkpointing
into a shared single file is most convenient. With such an approach, the size
of writes are often small and not aligned with file system boundaries.
Unfortunately for these applications, this preferred data layout results in
pathologically poor performance from the underlying file system which is
optimized for large, aligned writes to non-shared files. To address this
fundamental mismatch, we have developed a virtual parallel log structured file
system, PLFS. PLFS remaps an application's preferred data layout into one which
is optimized for the underlying file system. Through testing on PanFS, Lustre,
and GPFS, we have seen that this layer of indirection and reorganization can
reduce checkpoint time by an order of magnitude for several important
benchmarks and real applications.  
} 
}

### 2009-07-29 SC09/eval.tex
attached is the most recent version of the paper.  I'll do another big push tomorrow and then I'll submit it since I'll be on vacation all next week.  I think I've addressed most of the reviewer's comments but here's a couple that maybe still need to be addressed (or ignored):
the addition of metadata benchmark such as mdtest would make paper
stronger

the paper does not mention about file locks.  file system semantics is
not clear.

there should be a discussion about file system approach vs library
approach.  library can speed up N-1 w/out sacrificing performance for
other ops like metadata q

The top three graphs, 
Figures~\ref{eval:panfs},~\ref{eval:gpfs},~and~\ref{eval:lustre}, present the
results of our study using the LANL synthetic checkpoint tool, \Term{MPI-IO
Test}~\cite{mpi-io-test}, on three different parallel file systems, PanFS, GPFS,
and Lustre. For each of these graphs, the size of each \syscall{write} was
47001 bytes (a small, unaligned number observed in actual applications to be
particularly problematic for file systems). \syscall{Writes} were
issued until two minutes had elapsed. Although this is atypical since 
applications tend to write a fixed amount of data instead of writing for
a fixed amount of time, we have observed that this allows representative 
bandwidth measurements with a predictable runtime.

There are several things to notice in these graphs. The first is that these
are the same three graphs that we presented in Figure~\ref{fig-motivation}
except that we have now added a third line to each. The three lines show the
bandwidth achieved by writing an N-N pattern directly to the \upfs, the
bandwidth achieved by writing an N-1 pattern directly to the \upfs, and the
third line is the bandwidth achieved by writing an N-1 pattern {\em indirectly}
to the \upfs\ through \plfs.

@@

Although \plfs\ does improve the bandwidth of N-1 patterns on GPFS and Lustre,
the improvement is not as large as it is on PanFS. This is because the scale
of the experiments on PanFS are 200 times larger than on the other two
platforms. In the inset in Figure~\ref{eval:panfs} at the extreme low values
for the number of processes, we see that \plfs\ does not scale N-1 bandwidth
as fast as N-N scales for PanFS as well. This is due to overheads incurred by
both FUSE and \plfs; these overheads limit the total bandwidth achieved by any
single compute node relative to N-N. For HPC systems at extreme scale this
limitation does not matter since aggregate bandwidth across multiple nodes is
more relevant than the bandwidth from a single node. In this case the real
limiting factor is the total bandwidth capacity of the storage fabric, which
generally saturates when using only a small fraction of compute nodes.
However, with poorly-behaved IO patterns (\ie\ N-1), even very large jobs may
not reach these bandwidth limits because they will be more severely
constrained by file system limitations as exemplified by
Figure~\ref{fig-motivation}. \plfs\ is designed to remove these file system
limitations, so that jobs can achieve higher bandwidth and reach the
same limitation of the storage fabric which is constraining N-N bandwidth.

An unusual feture of the inset graph in Figure~\ref{eval:panfs} is that there
are localized bottlenecks about 700 processes.  This is due to the
architecture of Roadrunner and it's connection to PanFS.  Roadrunner is split
into seventeen sub-clusters, \Term{CUs}, each of which can run 720 proceses.
The CUs are fully connected via an Infiniband fat tree, however, in order to
minimize network infrastructure costs, the storage system is partitioned into
multiple sub-nets. Each CU currently has six specialized IO nodes, one for
each sub-net; these six IO nodes are the storage bandwidth bottleneck.
Therefore, as a job grows within a CU, it quickly saturates the IO node
bandwidth; when it grows into another CU, its bandwidth increases sharply due
to gaining six additional IO nodes: this explains the "stair-step" behavior of
the N-N and \plfs\ lines in this graph.

We do see this same behavior in our GPFS and Lustre graphs, but due to the
very small size of our test systems, we do not have enough compute nodes to
saturate the available storage bandwidth and thus neither the N-N nor the
\plfs\ lines in these graphs reach a storage bandwidth bottleneck. We will
examine the FUSE and PLFS overheads in greater detail in
Section~\ref{overhead}.

@@

The second row in Figure~\ref{fig-eval}, consisting of
Figures~\ref{eval:lanl1},~\ref{eval:lanl2},~and~\ref{eval:lanl3}, shows the
results of using \plfs\ to improve the bandwidth of three important LANL
applications which use N-1 checkpoint patterns. The application
shown in Figure~\ref{eval:lanl1} is the application whose developers augmented
their checkpoint routine with a new checkpoint library called bulkio, which
aggregates and buffers writes into larger, contiguous writes more friendly to
the \upfs. Therefore we show four lines in this graph: one for the old
checkpoint method which uses MPI-IO routines to directly write data to PanFS, a
line for the bulkio method writing directly to PanFS, and a second line for
each of these techniques writing indirectly to PanFS via \plfs. Notice that
\plfs\ is not able to improve the bulkio technique which loses a significant
amount of bandwidth due to aggregating data and reducing parallelism by
reducing the number of writers. However, when the older MPI-IO technique is
used with \plfs, \plfs\ is able to exploit the additional parallelism possible
by using all of the processes as writers. This data was collected on a small,
Opteron-only, \rrz. The next two graphs show similar results; using \plfs\ to
rearrange an N-1 pattern yields significantly improved bandwidths.
Figure~\ref{eval:lanl2} was also run on the \rrz, whereas
Figure~\ref{eval:lanl3} was run on Roadrunner itself; this can be seen by the
much larger scale (on both the \yaxis\ and the \xaxis) for
Figure~\ref{eval:lanl3}. Note that hese are extremely important applications
at LANL; it has been estimated that they account for more than half of the
total computing cycles at LANL. Although these applications must remain
anonymous, traces of their IO are available~\cite{plfs-maps}.

@@

The remainder of the graphs show various other parallel IO benchmarks with an
N-1 checkpoint pattern. QCD~\cite{qcd-io} in Figure~\ref{eval:qcd} shows a
large improvement using \plfs\ but this improvement actually {\em degrades} as
the number of processes increases. This is because the amount of data written
by QCD is a small constant value and the overhead due to container creation
incurred in the \syscall{open} becomes proportionally greater as the amount of
data written by each process decreases. BTIO~\cite{btio-io} in
Figure~\ref{eval:btio} also shows a large improvement but we currently only
have one data point; we will have more in the final version.
FLASH~IO~\cite{flash-io} and Chombo~IO\cite{chombo-io}, shown respectively in
Figures~\ref{eval:flash}~and~\ref{eval:chombo}, both show improvement due to
\plfs\ which scales nicely with an increasing number of processes. FLASH was
run on Roadrunner whereas Chombo was run on the \rrz; both were built with the
HDF5 library~\cite{hdf5}.

@@

[Note to the reviewer: Only FLASH and LANL3 were run at large-scale using both
\plfs\ and the \upfs. We will have fuller results for the final experiment and
we expect at least two orders of magnitude improvement for all of them except
QCD.]

@@

We measured the time to read these files using three different read
scenarios. One emulates a \Term{uniform restart} in which the number of
processes resuming the application is the same as the number that wrote the
last checkpoint. In this case, each rank within the parallel job reads the
same data in the same order as the previous writer of that same rank. In
contrast, a \Term{non-uniform restart} is one in which the number of
readers is different from the number of writers and the read offsets are
therefore not aligned with the previous writes. In this case, we emulate the
scenario where an application resumes after a failure by simply running on the
newly reduced number of nodes. The number of reads is not affected by the size
of the job for typical N-1 checkpoints. Each read is extracting a region for a
particular variable within the simulation. The size of this region depends on
the number of processes within the job. Each gets $1/Nth$ of the region.
Specifically, for a non-uniform restart in which the number of processes, $N$,
has been reduced by $M$, the size of each read will be $N/(N-M)$ times the size
of the original writes. The third read scenario we emulated is a typical
scenario in which a relatively, small fixed number of processes reads a
checkpoint in order to save it onto an archive system. For these measurements,
we used LANL's archive copy utility with fours readers each running on their
own node; each reader read just one contiguous 5~\GB\ region by issuing 
sequential 1~\MB\ reads.

>>>>

The top three graphs, 
Figures~\ref{eval:panfs},~\ref{eval:gpfs},~and~\ref{eval:lustre}, present the
results of our study using the LANL synthetic checkpoint tool, \Term{MPI-IO
Test}~\cite{mpi-io-test}, on three different parallel file systems, PanFS, GPFS,
and Lustre. 
There are several things to notice in these graphs. The first is that these
are the same three graphs that we presented in Figure~\ref{fig-motivation}
except that we have now added a third line to each. The three lines show the
bandwidth achieved by writing an N-N pattern directly to the \upfs, the
bandwidth achieved by writing an N-1 pattern directly to the \upfs, and the
third line is the bandwidth achieved by writing an N-1 pattern {\em indirectly}
to the \upfs\ through \plfs.

@@

Although \plfs\ does improve the bandwidth of N-1 patterns on GPFS and Lustre,
the improvement is not as large as it is on PanFS. This is because the scale of
the experiments on PanFS are 200 times larger than on the other two platforms.
In the inset in Figure~\ref{eval:panfs} at the extreme low values for the
number of processes, we see that \plfs\ does not scale N-1 bandwidth as fast as
N-N scales for PanFS as well. This is due to overheads incurred by both FUSE
and \plfs; these overheads limit the total bandwidth achieved by any single
compute node relative to N-N. For HPC systems at extreme scale this limitation
does not matter since aggregate bandwidth across multiple nodes is more
relevant than the bandwidth from a single node. In this case the real limiting
factor is the total bandwidth capacity of the storage fabric, which generally
saturates when using only a small fraction of compute nodes.  However, with
poorly-behaved IO patterns (\ie\ N-1), even very large jobs may not reach these
bandwidth limits because they will be more severely constrained by file system
limitations as exemplified by Figure~\ref{fig-motivation}. \plfs\ is designed
to remove these file system limitations, so that jobs can achieve higher
bandwidth and reach the same limitation of the storage fabric which is
constraining N-N bandwidth.  We will examine the FUSE and PLFS overheads in
greater detail in Section~\ref{overhead}.

An unusual feature of the inset graph in Figure~\ref{eval:panfs} is that there
are localized bottlenecks about 700 processes.  This is due to the architecture
of Roadrunner and it's connection to its 1000 blade PanFS storage system.
Roadrunner is split into seventeen sub-clusters, \Term{CUs}, each of which can
run 720 proceses.  The CUs are fully connected via an Infiniband fat tree,
however, in order to minimize network infrastructure costs, the storage system
is partitioned into multiple sub-nets. Each CU currently has six specialized IO
nodes, one for each sub-net; these six IO nodes are the storage bandwidth
bottleneck.  Therefore, as a job grows within a CU, it quickly saturates the IO
node bandwidth; when it grows into another CU, its bandwidth increases sharply
due to gaining six additional IO nodes: this explains the ``stair-step"
behavior of the N-N and \plfs\ lines in this graph.

We do see this same behavior in our GPFS and Lustre graphs, but due to the
very small size of our test systems, we do not have enough compute nodes to
saturate the available storage bandwidth and thus neither the N-N nor the
\plfs\ lines in these graphs reach a storage bandwidth bottleneck.

@@

The second row in Figure~\ref{fig-eval} shows the results of using \plfs\ to
improve the bandwidth of three important LANL applications which use N-1
checkpoint patterns. The application shown in Figure~\ref{eval:lanl1} is the
application whose developers augmented their checkpoint routine with a new
checkpoint library called bulkio, which aggregates and buffers writes into
larger, contiguous writes more friendly to the \upfs. Therefore we compare the
bulkio method writing directly to PanFS to the old MPI-IO method writing
indirectly to PanFS via \plfs.  This data was collected on Roadrunner.  The
next two graphs show similar results; using \plfs\ to rearrange an N-1 pattern
yields significantly improved bandwidths.  Figure~\ref{eval:lanl2} was run on a
small \rrz\ which consists of 128 8-core compute nodes with a 66 blade PanFS
storage system, whereas Figure~\ref{eval:lanl3} was run on Roadrunner itself.
Note that these are extremely important applications at LANL; it has been
estimated that they account for more than half of the total computing cycles at
LANL.  Although these applications must remain anonymous, traces of their IO
are available~\cite{plfs-maps}.

@@

The remainder of the graphs show various other parallel IO benchmarks with an
N-1 checkpoint pattern. QCD~\cite{qcd-io} in Figure~\ref{eval:qcd} shows a
large improvement using \plfs\ but this improvement actually {\em degrades} as
the number of processes increases. This is because the amount of data written
by QCD is a small constant value and the overhead due to container creation
incurred in the \syscall{open} becomes proportionally greater as the amount of
data written by each process decreases. BTIO~\cite{btio-io} in
Figure~\ref{eval:btio} also shows a large improvement but only
for one data point.
FLASH~IO~\cite{flash-io} and Chombo~IO\cite{chombo-io}, shown respectively in
Figures~\ref{eval:flash}~and~\ref{eval:chombo}, both show improvement due to
\plfs\ which scales nicely with an increasing number of processes. FLASH was
run on Roadrunner whereas Chombo was run on the \rrz; both were built with the
HDF5 library~\cite{hdf5}.

@@

@@

We measured the time to read these files using three different read
scenarios. One emulates a \Term{uniform restart} in which the number of
processes resuming the application is the same as the number that wrote the
last checkpoint. In this case, each process within the parallel job reads the
same data in the same order as one of the previous writers. In
contrast, a \Term{non-uniform restart} is one in which the number of
readers is different from the number of writers and the read offsets are
therefore not aligned with the previous writes. In this case, we emulate the
scenario where an application resumes after a failure by simply running on the
newly reduced number of nodes. The number of reads is not affected by the size
of the job for typical N-1 checkpoints. Each read is extracting a region for a
particular variable within the simulation. The size of this region depends on
the number of processes within the job. Each gets $1/Nth$ of the region.
Specifically, for a non-uniform restart in which the number of processes, $N$,
has been reduced by $M$, the size of each read will be $N/(N-M)$ times the size
of the original writes. The third read scenario we emulated is a typical
scenario in which a relatively, small fixed number of processes reads a
checkpoint in order to save it onto an archive system. For these measurements,
we used LANL's archive copy utility with fours readers each running on their
own node; each reader read just one contiguous 5~\GB\ region by issuing 
sequential 1~\MB\ reads.

### 2009-07-29 SC09/future.tex
attached is the most recent version of the paper.  I'll do another big push tomorrow and then I'll submit it since I'll be on vacation all next week.  I think I've addressed most of the reviewer's comments but here's a couple that maybe still need to be addressed (or ignored):
the addition of metadata benchmark such as mdtest would make paper
stronger

the paper does not mention about file locks.  file system semantics is
not clear.

there should be a discussion about file system approach vs library
approach.  library can speed up N-1 w/out sacrificing performance for
other ops like metadata q

\scribble{I removed the 'cute' part of this. Sorry, john}
We initially developed \plfs\ in order to test our hypothesis that an
interposition layer could rearrange checkpoint patterns such that the
convenience of N-1 patterns could be preserved while achieving the bandwidth of
N-N.  However, after achieving such large improvements with real LANL
applications, we were compelled to harden it 
into a production file system. Consisting of about three thousand lines of
code, \plfs\ is currently a permanent mount point on Roadrunner where several
applications have begun evaluating whether it is appropriate for their write 
patterns.

@@

Although \plfs\ works transparently with unmodified applications, in practice
some applications may need to make minor changes.  Applications that open
\plfs\ files in read-write mode will find that reading from these files can be
very slow.  As we discussed in Section~\ref{arch-read}, reads in \plfs\ are
handled by reading and aggregating the index files.  When files are opened in
write-read mode, this process must be repeated for every \syscall{read} since
intervening \syscall{writes} may have occurred.  Although the vast majority of
HPC applications are able to read in read-only mode, we do plan to investigate
removing this limitation in the future for those few applications who must occasionally
read while writing.  One possibility is to introduce metadata
servers which can synchronously maintain an aggregated index in memory.

>>>>

\scribble{I removed the 'cute' part of this. Sorry, john}
We initially developed \plfs\ in order to test our hypothesis that an
interposition layer could rearrange checkpoint patterns such that the
convenience of N-1 patterns could be preserved while achieving the bandwidth of
N-N.  However, after achieving such large improvements with real LANL
applications, we were compelled to harden it 
into a production file system. Consisting of about three thousand lines of
code, \plfs\ is currently an optional mount point on Roadrunner where several
applications have begun using it to improve their N-1 checkpoint performance.

@@

Although \plfs\ works transparently with unmodified applications, in practice
some applications may need to make minor changes.  Many users find it most
convenient to stage their input data sets and their executable code in the same
directory where they write their checkpoint files.  Since \plfs\ is
specifically a checkpoint file system and is not a general purpose file system,
this usage pattern may suffer a performance hit.  These users should instead
stage their input data sets and their executable code in a general purpose file
system and merely direct their N-1 checkpoint files to a \plfs\ directory.

Also, applications that open \plfs\ files in read-write mode will find that
reading from these files can be very slow.  As we discussed in
Section~\ref{arch-read}, reads in \plfs\ are handled by reading and aggregating
the index files.  When files are opened in write-read mode, this process must
be repeated for every \syscall{read} since intervening \syscall{writes} may
have occurred.  Although the vast majority of HPC applications are able to read
in read-only mode, we do plan to investigate removing this limitation in the
future for those few applications who must occasionally read while writing.
One possibility is to introduce metadata servers which can synchronously
maintain an aggregated index in memory.

### 2009-07-29 SC09/intro.tex
got the graphs looking good. got the abstract right. got the keywords, crap, authors, copyright.
git-svn-id: https://plfs.svn.sourceforge.net/svnroot/plfs/doc@7 979fb63a-54b4-4514-8d67-9ab8ea1b59df
attached is the most recent version of the paper.  I'll do another big push tomorrow and then I'll submit it since I'll be on vacation all next week.  I think I've addressed most of the reviewer's comments but here's a couple that maybe still need to be addressed (or ignored):
the addition of metadata benchmark such as mdtest would make paper
stronger

the paper does not mention about file locks.  f

\scribble{Need to define "its class" --Milo}
% computers are large, they have failures
In June 2008, Los Alamos National Labs (LANL), in partnership with IBM, broke
the petaflop barrier and claimed the top spot on the Top 500 list~\cite{top500}
with the Roadrunner supercomputer~\cite{roadrunner}.  Due to its unique hybrid
architecture, Roadrunner has only 3060 nodes, which is a relatively small
number for a supercomputer of its class.  Unfortunately, even at this size,
component failures are a frequent event.  These failures are particularly
problematic as many applications at LANL, and other High Performance Computing
(HPC) sites, have long run times in excess of days, weeks, and even months.

>>>>

\scribble{Need to define "its class" --Milo}
% computers are large, they have failures
In June 2008, Los Alamos National Labs (LANL), in partnership with IBM, broke
the petaflop barrier and claimed the top spot on the Top 500 list~\cite{top500}
with the Roadrunner supercomputer~\cite{roadrunner}.  Due to its unique hybrid
architecture, Roadrunner has only 3060 nodes, which is a relatively small
number for a supercomputer of its class.  Unfortunately, even at this size,
component failures are a frequent event.  For large parallel simulations at
LANL, and other High Performance Computing (HPC) sites, which can run for
several months, restarting a calculation from the beginning is not viable,
since the mean time to interrupt for today's supercomputers is measured in days
or hours and not months~\cite{google-failure,schroeder-failure}.

\input{patterns-fig}

### 2009-07-29 SC09/motivation.tex
attached is the most recent version of the paper.  I'll do another big push tomorrow and then I'll submit it since I'll be on vacation all next week.  I think I've addressed most of the reviewer's comments but here's a couple that maybe still need to be addressed (or ignored):
the addition of metadata benchmark such as mdtest would make paper
stronger

the paper does not mention about file locks.  file system semantics is
not clear.

there should be a discussion about file system approach vs library
approach.  library can speed up N-1 w/out sacrificing performance for
other ops like metadata q

% Why do we care about checkpointing
For large parallel simulations which can run for several months, restarting a
calculation from the beginning is not viable, since the mean time to interrupt
for today's supercomputers is measured in days or hours and not
months~\cite{google-failure,schroeder-failure}.  As a result, applications 
periodically save
the current state of their computation to persistent storage.  Later, after a
failure is encountered, the application can restart from the last saved
checkpoint.

@@

\input{patterns-fig}
\scribble{Should probably explain or remove 'false sharing' in figure}
\scribble{If those Argonne graphs were decipherable in B-W, I'd like to include
two of them to visualize the difference between strided and non.}

@@

Figure~\ref{fig-motivation} presents experimental data validating this discrepancy: An N-1 checkpoint pattern receives
only a small fraction of the bandwidth achieved by an N-N pattern on PanFS,
GPFS, and Lustre and does not scale with increased numbers of nodes.  The 
PanFS experiment, run on LANL's Roadrunner
supercomputer, shows a maximum observed N-N bandwidth of 31~\GBs\ compared
to a maximum observed N-1 bandwidth of less than 100~\MBs. 
Although we show PanFS results
using their default RAID-5 configuration, PanFS has also a RAID-10
configuration which reduces the implicit sharing caused by N-1 patterns when
two writers both need to update the same parity.  While this solution improves
scaling and offers much higher N-1 write bandwidth without sacrificing
reliability, it does so by writing every byte twice, a scheme that, at best,
can achieve only approximately half of the write bandwidth of N-N on RAID-5.
\plfs, however, as will be shown in Section~\ref{eval}, can get much closer.

The GPFS and Lustre experiments were run on much smaller systems. The GPFS
experiment was run using an archive attached to Roadrunner using its 
nine, quad-core, file transfer nodes.  The Lustre experiment was run using five
client machines, each with eight cores, and twelve Lustre servers.  All three
file systems exhibit similar behavior; N-N bandwidths are consistently higher
than N-1 by at least an order of magnitude.

>>>>

% Why do we care about checkpointing
\if 0
As a result, applications 
periodically save
the current state of their computation to persistent storage.  Later, after a
failure is encountered, the application can restart from the last saved
checkpoint. 
\fi

@@

\scribble{Should probably explain or remove 'false sharing' in figure}
\scribble{If those Argonne graphs were decipherable in B-W, I'd like to include
two of them to visualize the difference between strided and non.}

@@

Figure~\ref{fig-motivation} presents experimental data validating this
discrepancy: An N-1 checkpoint pattern receives only a small fraction of the
bandwidth achieved by an N-N pattern on PanFS, GPFS, and Lustre and does not
scale with increased numbers of nodes.  The PanFS experiment, run on LANL's
Roadrunner supercomputer using its 1000 blade PanFS storage system, shows a
maximum observed N-N bandwidth of 31~\GBs\ compared to a maximum observed N-1
bandwidth of less than 100~\MBs.  Although we show PanFS results using their
default RAID-5 configuration, PanFS has also a RAID-10 configuration which
reduces the implicit sharing caused by N-1 patterns when two writers both need
to update the same parity.  While this solution improves scaling and offers
much higher N-1 write bandwidth without sacrificing reliability, it does so by
writing every byte twice, a scheme that, at best, can achieve only
approximately half of the write bandwidth of N-N on RAID-5.  \plfs, however, as
will be shown in Section~\ref{eval}, can get much closer.

The GPFS and Lustre experiments were run on much smaller systems. The GPFS
experiment was run using an archive attached to Roadrunner using its 
nine, quad-core, file transfer nodes.  The Lustre experiment was run using five
client machines, each with eight cores, and twelve Lustre servers.  All three
file systems exhibit similar behavior; N-N bandwidths are consistently higher
than N-1 by at least an order of magnitude.  Measurements were gathered using
the LANL synthetic checkpoint tool, \Term{MPI-IO Test}~\cite{mpi-io-test}.
For each of these graphs, the size of each \syscall{write} was
47001 bytes (a small, unaligned number observed in actual applications to be
particularly problematic for file systems). \syscall{Writes} were
issued until two minutes had elapsed. Although this is atypical since 
applications tend to write a fixed amount of data instead of writing for
a fixed amount of time, we have observed that this allows representative 
bandwidth measurements with a predictable runtime.

### 2009-07-29 SC09/related-table.tex
git-svn-id: https://plfs.svn.sourceforge.net/svnroot/plfs/doc@5 979fb63a-54b4-4514-8d67-9ab8ea1b59df
attached is the most recent version of the paper.  I'll do another big push tomorrow and then I'll submit it since I'll be on vacation all next week.  I think I've addressed most of the reviewer's comments but here's a couple that maybe still need to be addressed (or ignored):
the addition of metadata benchmark such as mdtest would make paper
stronger

the paper does not mention about file locks.  file system semantics is
not clear.

there should be a discussion about file system approach vs li

\newcommand{\C}[1]{\centering{#1}}
%\newcommand{\C}[1]{\begin{center}#1\end{center}}
\hyphenation{Immediately}
\begin{table}
\begin{center}
\begin{tabular}{|r|p{19mm}|p{20mm}|p{23mm}|p{15mm}|p{19mm}| p{19mm}|}
\hline 
{\em } & 
\C{{\em Interposition Technique Used}} & 
\C{{\em No Extra Resources Used During}} & 
\C{{\em No Extra Resources Used After}} & 
\C{{\em Maintains Logical Format}} & 
\C{{\em Works with Unmodified Applications}} & 
\C{{\em Data Immediately Available}} 
\tabularnewline \hline
ADIOS & \C{Library} & \C{Yes} & \C{Yes} & \C{Yes} &  \C{No} & \C{Yes}
\tabularnewline \hline
{\tt stdchk} & \C{FUSE} & \C{No (LD, M)} & \C{No (LD, N, M)} & \C{Yes} & \C{Yes} & \C{Yes}
\tabularnewline \hline
%Neighbor &  \C{FUSE} &\C{No (M)} & \C{No (M, N)} & \C{Yes} & \C{Yes} & \C{Yes} 
%\tabularnewline \hline
Diskless &  \C{Library} &\C{No (M)} & \C{No (M)} & \C{No} & \C{No} & \C{Yes} 
\tabularnewline \hline
ZEST &  \C{FUSE} &\C{No (RD)} & \C{No (RD)} & \C{No} & \C{No} & \C{No} 
\tabularnewline \hline
PLFS & \C{FUSE} & \C{Yes} & \C{Yes} & \C{Yes} & \C{Yes} & \C{Yes} 
\tabularnewline \hline
\end{tabular}
\mycaption{tab-feature}{Techniques for improving N-1 Checkpointing}{
This table presents a comparison of the various techniques for reducing
N-1 checkpoint times.  For exposition, we have used various
shorthands: 
%Neighbor for Neighbor's Memory,
Diskless for Diskless 
Checkpointing, LD for local disk on the compute nodes, RD for remote disk 
on the storage system, M for memory, and N for network.  
}
\end{center}
\end{table}

>>>>

\newcommand{\C}[1]{\centering{#1}}
%\newcommand{\C}[1]{\begin{center}#1\end{center}}
\hyphenation{Immediately}
\begin{table*}
\begin{center}
\begin{tabular}{|r|p{19mm}|p{20mm}|p{23mm}|p{15mm}|p{19mm}| p{19mm}|}
\hline 
{\em } & 
\C{{\em Interposition Technique Used}} & 
\C{{\em No Extra Resources Used During}} & 
\C{{\em No Extra Resources Used After}} & 
\C{{\em Maintains Logical Format}} & 
\C{{\em Works with Unmodified Applications}} & 
\C{{\em Data Immediately Available}} 
\tabularnewline \hline
ADIOS & \C{Library} & \C{Yes} & \C{Yes} & \C{Yes} &  \C{No} & \C{Yes}
\tabularnewline \hline
{\tt stdchk} & \C{FUSE} & \C{No (LD, M)} & \C{No (LD, N, M)} & \C{Yes} & \C{Yes} & \C{Yes}
\tabularnewline \hline
%Neighbor &  \C{FUSE} &\C{No (M)} & \C{No (M, N)} & \C{Yes} & \C{Yes} & \C{Yes} 
%\tabularnewline \hline
Diskless &  \C{Library} &\C{No (M)} & \C{No (M)} & \C{No} & \C{No} & \C{Yes} 
\tabularnewline \hline
Sp Writ &  \C{Library} &\C{Yes} & \C{Yes} & \C{Yes} & \C{No} & \C{No} 
\tabularnewline \hline
ZEST &  \C{FUSE} &\C{No (RD)} & \C{No (RD)} & \C{No} & \C{No} & \C{No} 
\tabularnewline \hline
PLFS & \C{FUSE} & \C{Yes} & \C{Yes} & \C{Yes} & \C{Yes} & \C{Yes} 
\tabularnewline \hline
\end{tabular}
\mycaption{tab-feature}{Techniques for improving N-1 Checkpointing}{
This table presents a comparison of the various techniques for reducing
N-1 checkpoint times.  For exposition, we have used various
shorthands: 
%Neighbor for Neighbor's Memory,
Diskless for Diskless 
Checkpointing, Sp Writ for Lustre Split Writing, LD for local disk on the compute nodes, RD for remote disk 
on the storage system, M for memory, and N for network.  
}
\end{center}
\end{table*}

### 2009-07-29 SC09/related.tex
attached is the most recent version of the paper.  I'll do another big push tomorrow and then I'll submit it since I'll be on vacation all next week.  I think I've addressed most of the reviewer's comments but here's a couple that maybe still need to be addressed (or ignored):
the addition of metadata benchmark such as mdtest would make paper
stronger

the paper does not mention about file locks.  file system semantics is
not clear.

there should be a discussion about file system approach vs library
approach.  library can speed up N-1 w/out sacrificing performance for
other ops like metadata q

The Adaptable IO System (ADIOS)~\cite{adios} developed by the Georgia Institute
of Technology and Oak Ridge National Laboratory provides a high-level IO API
that can be used in place of HDF5 to do much more aggressive write-behind and
log-like reordering of data location within the checkpoint. While this technique
requires application modification, it enables interoperability with other
middleware libraries.

>>>>

The Adaptable IO System (ADIOS)~\cite{adios} developed by the Georgia Institute
of Technology and Oak Ridge National Laboratory provides a high-level IO API
that can be used in place of HDF5 to do much more aggressive write-behind and
log-like reordering of data location within the checkpoint. While this technique
requires application modification, it enables interoperability with other
middleware libraries.  Similarly, Lustre Split Writing~\cite{lsw}, uses a 
library approach and leverages Lustre's file joining mechanism to decouple
concurrent access at runtime as does \plfs.  However, Lustre Split Writing is
wed to the Lustre file system, requires application modification, and prevents
the data from being immediately available following a checkpoint.
